---
title: "poster2023"
format: html
editor: visual
---

Graphs to add for poster

# Topic Graphs

First go to `LDA.qmd` using the `Partnership_data.xlsx` data set and run everything up until the end of LDA model

Once you do that, this will be the graph for visualizing words in topics

```{r}
beta <- tpw.df %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>%
  ungroup() %>%
  arrange(topic, -beta)
```

```{r}
beta %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  scale_y_reordered() +
  xlab("Probability") +
  ylab("Word") + 
  labs(title = "Grouped Terms") +
  theme_minimal()
```

Saving graph image

```{r}
ggsave("beta.png")
```

Now graph for visualizing probability topics in activity description

```{r}
gamma <- data.frame(dpt.df)
gamma$document <- rep(1:dim(partnership_dfm) [1],5)
```

```{r}
g <- ggplot(gamma, aes(document, gamma, group = factor(topic), fill = factor(topic))) +
  geom_bar(stat = "identity") +
  facet_wrap(~ factor(topic), ncol = 1) +
  labs(title = "Topics of Each Document") +
  ylab("Probability") +
  xlab("Activity Description")
  theme_minimal()
```

Saving graph image

```{r}
ggsave("gamma.png")
```

```{r}
head(partnership_topic)
filter(partner_topic, quarter == 1 ) # 294
filter(partner_topic, quarter == 2) # 344
filter(partner_topic, quarter == 3) # 344
filter(partner_topic, quarter == 4) # 3
```

# Random Forest Importance Graph

Load `partnership_topic` and `gcf20_21` and run the following code chunks

```{r}
partnership_topic <- partnership_topic %>% 
  rename(shortcode = y2n, 
         level = Level..1.lowest.3.highest...n0...excluded.for.not.enough.info,
         contract_year = year,
         topic_1 = "1",
         topic_2 = "2",
         topic_3 = "3",
         topic_4 = "4",
         topic_5 = "5")
```

```{r}
gcf20_21 <- gcf20_21 %>%
  mutate(quarter = case_when(
    month %in% c(10,11,12) ~ 1,
    month %in% c(1,2,3) ~ 2,
    month %in% c(4,5,6) ~ 3,
    month %in% c(7,8,9) ~ 4
  )) %>%
  select(c("shortcode", "prescreens", "totalapps", "contract_year", "quarter"))
head(gcf20_21)
```

```{r}
gcf20_21 <- gcf20_21 %>%
  group_by(shortcode, quarter, contract_year) %>%
  summarise(across(where(is.numeric), sum))
head(gcf20_21)
```

```{r}
lda_gcf <- left_join(partnership_topic, gcf20_21, by = c("shortcode", "contract_year", "quarter"))
head(lda_gcf)
```

```{r}
library(randomForest)
```

Removing these columns for importance graph

```{r}
lda_gcf <- select(lda_gcf, -c("contract_year", "descrip", "document"))
lda_gcf$quarter <- as.factor(lda_gcf$quarter)
lda_gcf$level <- as.factor(lda_gcf$level)
head(lda_gcf)
```

```{r}
tra.tes <- sample(c(TRUE, FALSE), nrow(lda_gcf), replace = TRUE, prob=c(.7,.3))
tra <- lda_gcf[tra.tes, ]
tes <- lda_gcf[-tra.tes, ]
tra <- tra[complete.cases(tra), ]
```

```{r}
set.seed(123)
rf <- randomForest(level ~ ., data = tra, proximity = TRUE, importance = TRUE)
rf
```

```{r}
importance <- as.data.frame(round(importance(rf), digits = 4))
mdg <- importance[, "MeanDecreaseGini"]
df <- data.frame(feature = row.names(importance), importance = mdg)
```

```{r}
ggplot(df, aes(x = feature, y = impportance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Important Variables to Predict Activity Priority", x = "Variable", y = "Importance") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

```{r}
ggsave("rf.png")
```
